Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-25 main.tex
Minor bug.
Rewrite Exercise and Implementation Section
Updated img.
Added new picture.
changes.
Improving bibtex

\author{
\IEEEauthorblockN{Jan Ciesko}
\IEEEauthorblockA{\textit{Computer Science Reseach Center} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
jciesko@sandia.gov}
\and
\IEEEauthorblockN{David Poliakoff}
\IEEEauthorblockA{\textit{Computer Science Reseach Center} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
dzpolia@sandia.gov}
\and
\IEEEauthorblockN{David S. Hollman}
\IEEEauthorblockA{\textit{Computer Science Reseach Center} \\
\textit{Sandia National Laboratories}\\
Livermore, CA, USA\\
dshallm@sandia.gov}
\and
\IEEEauthorblockN{Author Name}
\IEEEauthorblockA{\textit{Computer Science Reseach Center} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
aname@sandia.gov}
\and
\IEEEauthorblockN{Author Name}
\IEEEauthorblockA{\textit{Computer Science Reseach Center} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
aname@sandia.gov}
\and
\IEEEauthorblockN{Author Name}
\IEEEauthorblockA{\textit{Computer Science Reseach Center} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
aname@sandia.gov}
}

@@

\input{src/1.Introduction}           %label: chap:introduction
\input{src/2.GenericProgramming.tex} %label: chap:background
\input{src/3.MachineModel.tex}             %label: chap:kokkosPM
\input{src/4.ProgammingModel.tex}     %label: chap:kokkosEMA
\input{src/5.Backend.tex}            %label: chap:kokkosBackend
\input{src/6.Exercises}              %label: chap:tutorial
\input{src/7.C++}                    %label: chap:c++
\input{src/8.Related}                %label: chap:related
\input{src/9.Conclusion}             %label: chap:conclusion

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\end{thebibliography}

\end{document}

>>>>

\author{
\IEEEauthorblockN{Jan Ciesko}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
jciesko@sandia.gov}
\and
\IEEEauthorblockN{David Poliakoff}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
dzpolia@sandia.gov}
\and
\IEEEauthorblockN{David S. Hollman}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Livermore, CA, USA\\
dshallm@sandia.gov}
\and
\IEEEauthorblockN{Christian C. Trott}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
ctrott@sandia.gov}
\and
\IEEEauthorblockN{Damien Lebrun-Grandié}
\IEEEauthorblockA{\textit{Computational Sciences and Engineering} \\
\textit{Oak Ridge National Laboratory}\\
Oak Ridge, TN, USA\\
lebrungrandt@ornl.gov}
}

@@

\input{src/1.Introduction}           %label: chap:introduction
\input{src/2.GenericProgramming}     %label: chap:background
\input{src/3.MachineModel}           %label: chap:kokkosPM
\input{src/4.ProgammingModel}        %label: chap:kokkosEMA
\input{src/5.Example}                %label: chap:example
\input{src/6.Backend}                %label: chap:backend
\input{src/8.Related}                %label: chap:related
\input{src/9.Conclusion}             %label: chap:conclusion

\bibliographystyle{IEEEtran}
\bibliography{ref}
\end{document}

### 2020-01-25 src/1.Introduction.tex
Improving bibtex

Kokkos\footnote{from Greek for grain or seed}~\cite{KOKKOS}, is a performance-portable, parallel programming model for C++ developed collaboratively between national laboratories as part of the US Department of Energy Exascale Project\cite{ECP}. It consists of a programming model specification and a library. The specification defines execution and memory model and an application programming interface (library API). The runtime library exposes programming primitives, namely, patterns and types and maps them to the underlying abstract machine model. The machine model creates a generic view of compute hardware and memories. The Kokkos library maps execution primitives to vendor-specific programming models, libraries, and memory layouts. Kokkos has been successfully used in many large-scale projects at SNL, University of Utah, Oak-Ridge National Laboratories, Los Alamos National Laboratories and others~\cite{KOKKOSUSECASE2}. We list its key properties as follows.

@@

The rest of this paper is structured as follows. The next Chapter gives an overview on parallel programming concepts and presents the notion of \emph{semantic capture}. Chapter~\ref{chap:kokkosMM} presents the abstract machine model used to derive required semantic information for performance portability in Kokkos. Chapters~\ref{chap:kokkosPM} and ~\ref{chap:kokkosExample} present the Kokkos programming model and an example application. In Chapter~\ref{chap:backend} provides insight into the Kokkos library implementation. Chapters~\ref{chap:c++} and~\ref{chap:related} show the alignment to C++ standardization efforts and discuss related work and lastly Chapter~\ref{chap:conclusion} concludes this work.

>>>>

Kokkos\footnote{from Greek for grain or seed}~\cite{KOKKOS}, is a performance-portable, parallel programming model for C++ developed collaboratively between national laboratories as part of the US Department of Energy Exascale Project\cite{ECP}. It consists of a programming model specification and a library. The specification defines execution and memory model and an application programming interface (library API). The runtime library exposes programming primitives, namely, patterns and types and maps them to the underlying abstract machine model. The machine model creates a generic view of compute hardware and memories. The Kokkos library maps execution primitives to vendor-specific programming models, libraries, and memory layouts. Kokkos has been successfully used in many large-scale projects at SNL, University of Utah, Oak-Ridge National Laboratories, Los Alamos National Laboratories and others~\cite{KOKKOSPROJS}. We list its key properties as follows.

@@

The rest of this paper is structured as follows. The next Chapter gives an overview on parallel programming concepts and presents the notion of \emph{semantic capture}. Chapter~\ref{chap:kokkosMM} presents the abstract machine model used to derive required semantic information for performance portability in Kokkos. Chapters~\ref{chap:kokkosPM} and ~\ref{chap:kokkosExample} present the Kokkos programming model and an example application. In Chapter~\ref{chap:backend} we provide insight into a Kokkos back-end implementation. Chapters~\ref{chap:related} discusses related work and lastly Chapter~\ref{chap:conclusion} concludes this work.

### 2020-01-25 src/8.Related.tex
Improving bibtex

\section{Related Work}\label{chap:related}
RAJA~\cite{RAJA} is a C++ abstraction library with similarities to Kokkos. The important different however is that Kokkos is descriptive, while RAJA is prescriptive. In Kokkos, the framework determines how an application is mapped to the underlying hardware. RAJA provides functionality that exposes hardware details and relies on the developer to try different strategies to map kernels to architectures. RAJA is a viable option for class education in which the educator would like to expose students to the semantics of different programming models without exposing them to the detailed syntax of those models.

DPC++\cite{DPCPP} is a model developed by Intel\textsuperscript{\textregistered} for expressing parallelism across Intel architectures. In addition to sharing a descriptive philosophy with Kokkos, it also shares parallel patterns. Unlike Kokkos, at time of publication it targets only Intel architectures and is not vendor-agnostic. A course targeting Intel architectures, DPC++ could be a valid option.

C++ increasingly supports parallel constructs. We have highlighted contributions of Kokkos to C++ in~\ref{chap:c++}. It is becoming increasingly possible to teach parallelism in C++ and relying on modern C++ compilers. However, at time of publication, C++ does not support accelerators, or many of the complexities of parallel computing related to accelerator programming. In courses that do not target accelerators, this can be a relevant option.

>>>>

\section{Related Work}\label{chap:related}

C++ increasingly supports parallel constructs. It is becoming increasingly possible to teach parallelism in C++ and relying on modern C++ compilers. However, at time of publication, C++ does not support accelerators, or many of the complexities of parallel computing related to accelerator programming. In courses that do not target accelerators, this can be a relevant option.

Over the past decade the Kokkos programming model has contributed to a number of current and near-future ISO-C++ features, including \emph{atomic\_ref}\cite{wg21_p0019}, \emph{mdspan}\cite{MDSPAN}, and executors.  In particular, the design of C++ executors, which are similar to Kokkos execution spaces, were influenced by this programming model. As the demand to extract performance from increasingly deep and increasingly asynchronous software stacks across a wide variety of domains increases, many C++ experts expect executors to become a central abstraction in any performance-oriented software stack.

RAJA~\cite{RAJA} is a C++ abstraction library with similarities to Kokkos. The important different however is that Kokkos is descriptive, while RAJA is prescriptive. In Kokkos, the programming models determines how an application is mapped to the underlying hardware. RAJA provides functionality that exposes hardware details and relies on the developer to try different strategies to map kernels to architectures. RAJA is a viable option for class education in which the educator would like to expose students to the semantics of different programming models without exposing them to the detailed syntax of those models.

DPC++\cite{DPCPP} is a model developed by Intel\textsuperscript{\textregistered} for expressing parallelism across Intel architectures. In addition to sharing a descriptive philosophy with Kokkos, it also shares parallel patterns. Unlike Kokkos, at time of publication it targets Intel architectures only and is not vendor-agnostic. A course targeting Intel architectures, DPC++ could be a valid option.
